---
title: "Route Profiles"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    logo: septa.png
    favicon: septa.png
    source_code: embed
runtime: shiny
---

```{r global, include=FALSE}
library(flexdashboard)
library(dplyr)
library(magrittr)
library(ggplot2)
library(gridExtra)
library(shiny)
library(leaflet)
library(geojsonio)
library(sp)
library(rgdal)
library(DT)
library(lubridate)
#Load Data
setwd("C:/Users/bndodson/Documents/GitHub/route-profiles")
Rt_16 <- read.csv("Route16.csv")
Rt_21 <- read.csv("Route21.csv")
Rt_42 <- read.csv("Route42.csv")
Characterization <- read.csv("Characterization.csv")

#Compile and Clean Data
Rt_16 <- subset(Rt_16, Service.Type=='WEEKDAY')
data <- rbind(Rt_21, Rt_42, Rt_16)
data$Route <- data$ï..Route
data <- data[complete.cases(data), ]

#Create sorting column
data$Sort[data$Period=='Early AM'] <- '1-Early AM'
data$Sort[data$Period=='AM Peak'] <- '2-AM Peak'
data$Sort[data$Period=='Midday'] <- '3-Midday'
data$Sort[data$Period=='PM Peak'] <- '4-PM Peak'
data$Sort[data$Period=='Evening'] <- '5-Evening'
data$Sort[data$Period=='Late Night'] <- '6-Late Night'

data$Duration <- hms(data$Avg..Time)
data$Duration <- (hour(data$Duration)*60)+minute(data$Duration)+(second(data$Duration)/60)

#Group Data by Route and Period, Round values
data <- data %>%
  group_by(Route, Sort) %>%
  summarise(mean_ons = mean(Avg..Ons), mean_offs = mean(Avg..Offs), mean_speed = mean(Avg..M.H), 
            max_load = max(Max.Load), mean_duration = mean(Duration), mean_pv = mean(Avg..P.Vh), 
            mean_pm = mean(Avg..P.M), mean_dist = mean(Avg..Dist), mean_pmiles = mean(Avg..PM))
data <- data %>% 
  mutate_if(is.numeric, round)
names(data)[names(data) == "Sort"] <- "Period"

plotdat <- subset(data, Route=='16')
```

Sidebar {.sidebar}
==================

```{r}
 selectInput("filterA", label = h4("Route:"), 
    choices = list("16" = 16, "21" = 21, "42" = 42), 
    selected = 16) 
 selectInput("filterB", label = h4("Service:"), 
    choices = list("Weekday" = "WEEKDAY", "Saturday" = "SATURDAY", "Sunday" = "SUNDAY"), 
    selected = "Weekday")
 selectInput("filterC", label = h4("Period:"), 
    choices = list("Fall 2019" = "Fall-19", "Spring 2019" = "Spring 2019"), 
    selected = "Fall 2019")
 
```
*Currently under development. For internal use only.*

*Right-click on a chart to save it as an image for use in a document or presentation.*

[Open Data](http://septaopendata-septa.opendata.arcgis.com/)

Ridership {data-icon="fa-user"}
==================

Row
-----------------------------

### Table: Total Ridership, Pass/Hr/Mi/Trip

```{r}
tabdat <- subset(plotdat, select = -c(Route, mean_ons, mean_offs, mean_speed, max_load, mean_duration))
names(tabdat)[names(tabdat) == "mean_pv"] <- "Mean Passengers per Vehicle"
names(tabdat)[names(tabdat) == "mean_pm"] <- "Mean Passengers per Mile"
names(tabdat)[names(tabdat) == "mean_dist"] <- "Mean Trip Distance"
names(tabdat)[names(tabdat) == "mean_pmiles"] <- "Mean Passenger Miles"
DT::datatable(tabdat, options = list(
  bPaginate = FALSE
))
```

### Boardings

```{r, results='hide'}
b <- 
  ggplot(data=plotdat, aes(x=Period, y=mean_ons, fill=Period)) +
  geom_bar(stat='identity') +  
  geom_text(aes(label=mean_ons), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("Average Boardings per Trip") +
  ylab("Boardings (Passengers)")
b
```

Row
----------------------------
### Peak Max Load

```{r, results='hide'}
c <- 
  ggplot(data=plotdat, aes(x=Period, y=max_load, fill=Period)) +
  geom_bar(stat='identity') +  
  geom_text(aes(label=mean_offs), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("Maximum Load of All Trips") +
  ylab("Maximum Average Load per Trip (Passengers)")
c
```

### Seat Utilization

```{r, results='hide'}
#d
```

Performance {data-icon="fa-clipboard-check"}
==================

Row
-----------------------------

### OTP: Early AM

```{r}
rate1 <- 83
gauge(rate1, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

### OTP: AM Peak

```{r}
rate2 <- 74
gauge(rate2, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

### OTP: Midday
```{r}
rate3 <- 91
gauge(rate3, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

### OTP: PM Peak
```{r}
rate4 <- 59
gauge(rate4, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

### OTP: Evening
```{r}
rate5 <- 81
gauge(rate5, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

### OTP: Late Night
```{r}
rate6 <- 89
gauge(rate6, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(85, 100), warning = c(65, 84), danger = c(0, 64)
))
```

Row
-----------------------------

### Speed

```{r, results='hide'}
g <- 
  ggplot(data=plotdat, aes(x=Period, y=mean_speed, fill=Period)) +
  geom_bar(stat='identity') +  
  geom_text(aes(label=mean_speed), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  ggtitle("Mean Speed") +
  theme_minimal() + 
  theme(legend.position = "none") +
  ylab("Speed (MPH)")
g
```

----------------------------
### Trip Duration

```{r, results='hide'}
h <- 
  ggplot(data=plotdat, aes(x=Period, y=mean_duration, fill=Period)) +
  geom_bar(stat='identity') +  
  geom_text(aes(label=mean_duration), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  ggtitle("Mean Trip Duration") +
  theme_minimal() + 
  theme(legend.position = "none") +
  ylab("Trip Duration (Minutes)")
h
```

Operations {data-icon="fa-bus"}
==================

Row
-----------------------------

### Table: Distance, Vehicle Miles, Peak Vehicles

```{r, results='hide'}
#i
```

### Passengers/Vehicle

```{r, results='hide'}
j <- 
  ggplot(data=plotdat, aes(x=Period, y=mean_pv, fill=Period)) +
  geom_bar(stat='identity') +  
  geom_text(aes(label=mean_pv), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle("Average Passengers per Vehicle") +
  ylab("Passengers")
j
```

Row
----------------------------
### Expenses/Revenue

```{r, results='hide'}
#k
```

### Peak Vehicles

```{r, results='hide'}
#l
```

Characterization {data-icon="fa-table"}
==================
```{r}
Characterization <- subset(Characterization, select = -c(ï..Route))
names(Characterization)[names(Characterization) == "Passenger.Comfort"] <- "Passenger Comfort"
names(Characterization)[names(Characterization) == "Running.Times"] <- "Running Times"
names(Characterization)[names(Characterization) == "Stop.Spacing"] <- "Stop Spacing"
char <- t(Characterization)
DT::datatable(char, options = list(
  bPaginate = FALSE
))
```

Map {data-icon="fa-map"}
==================
```{r, message=FALSE, warning=FALSE}
routelines <- rgdal::readOGR("https://opendata.arcgis.com/datasets/cc102b601bb947c8bb947937c80e3aa5_0.geojson", verbose='FALSE')
ridership <- rgdal::readOGR("https://opendata.arcgis.com/datasets/e09e9f98bdf04eada214d2217f3adbf1_0.geojson", verbose='FALSE')

maproute <- subset(routelines, Route==16)
ridership_16 <- subset(ridership, Route==16)
ridership_16N <- subset(ridership_16, Direction == "Northbound")
ridership_16S <- subset(ridership_16, Direction == "Southbound")


m = leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron, group =  "Positron") %>%
  addPolylines(data = maproute,
               weight = 3, 
               group = "Route 16") %>%
  addCircles(data = ridership_16N,
             weight = 1, 
             color = '#8000ff',
             radius = ridership_16N$Weekday_Boards+ridership_16N$Weekday_Leaves, 
             popup = paste("Stop:", ridership_16N$Stop_Name, "<br>",
                           "Direction:", ridership_16N$Direction, "<br>",
                           "Weekday Boards:", ridership_16N$Weekday_Boards,"<br>",
                           "Weekday Leaves:", ridership_16N$Weekday_Leaves, "<br>"),
             group = "Ridership Northbound"
             )%>%
  addCircles(data = ridership_16S,
             weight = 1, 
             color = 'royalblue',
             radius = ridership_16S$Weekday_Boards+ridership_16S$Weekday_Leaves,
             popup = paste("Stop:", ridership_16S$Stop_Name, "<br>",
                           "Direction:", ridership_16S$Direction, "<br>",
                           "Weekday Boards:", ridership_16S$Weekday_Boards, "<br>",
                           "Weekday Leaves:", ridership_16S$Weekday_Leaves, "<br>"), 
             group = "Ridership Southbound")%>%
  addLayersControl(
    overlayGroups = c("Route 16", "Ridership Northbound", "Ridership Southbound"),
    options = layersControlOptions(collapsed = FALSE)
    )%>%
  addLegend(
    position = "bottomright",
    colors = c("royalblue", "#8000ff"), 
    labels = c("Ridership Northbound", "Ridership Southbound")
  )

  
m
```
