---
title: "Route Profiles"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    logo: septa.png
    favicon: septa.png
    source_code: embed
runtime: shiny
---

```{r global, message=FALSE, warning=FALSE, include=FALSE}
library(flexdashboard)
library(dplyr)
library(ggplot2)
library(shiny)
library(leaflet)
library(rsconnect)
library(readr)
library(plotly)
library(DT)

OTP <- read_csv("data/OTP.csv")
OTP[OTP == 0] <- NaN
OTP <- OTP[complete.cases(OTP), ]
Characterization <- read_csv("data/Characterization.csv")
data <- read_csv("data/data.csv")
OTP_hist <- read_csv("data/OTP_hist_long.csv")
opstats <- read_csv("data/opstats.csv")
routelines <- rgdal::readOGR("https://opendata.arcgis.com/datasets/cc102b601bb947c8bb947937c80e3aa5_0.geojson", verbose='FALSE')
ridership <- rgdal::readOGR("https://opendata.arcgis.com/datasets/e09e9f98bdf04eada214d2217f3adbf1_0.geojson", verbose='FALSE')

```

Sidebar {.sidebar}
==================

```{r message=FALSE, warning=FALSE}
 selectInput("filterA", label = h4("Route:"), 
    choices = list("16" = 16, "21" = 21, "23" = 23, "42" = 42, "310" = 310, "LUCY Gold" = "LUCY GO"), 
    selected = 16) 
```
*Currently under development. For internal use only.*

*Right-click on a chart to save it as an image for use in a document or presentation.*

[Open Data](http://septaopendata-septa.opendata.arcgis.com/)

*Note: OTP data not availible for all routes at this time.*
```{r message=FALSE, warning=FALSE}
route_selection <- '42'
plotdat <- subset(data, Route==route_selection)
OTP_plot <- subset(OTP, Route==route_selection)


```

Ridership {data-icon="fa-user"}
==================


Row
-----------------------------

### Table: Total Ridership, Pass/Hr/Mi/Trip

```{r message=FALSE, warning=FALSE}
renderDT({
  datat1 <- subset(data, Route==input$filterA)
  datat1 <- subset(datat1, select = -c(X1, Route, mean_ons, mean_mph, max_load, mean_duration))
  names(datat1)[names(datat1) == "mean_ppvh"] <- "Mean Pass. per Vehicle"
  names(datat1)[names(datat1) == "mean_ppm"] <- "Mean Pass. per Mile"
  names(datat1)[names(datat1) == "mean_dist"] <- "Mean Trip Distance"
  names(datat1)[names(datat1) == "mean_pm"] <- "Mean Pass. Miles"
  DT::datatable(datat1, options = list(
    bPaginate = FALSE))
})
```

### Mean Boardings and Maximum Load

```{r message=FALSE, warning=FALSE}
renderPlot({
  bdat <- subset(data, Route==input$filterA)
  titleb <- sprintf("Average Boardings (Bars) + Max Load (Line) - Route %s", input$filterA)
  b <- ggplot(data=bdat, aes(x=Period, y=mean_ons, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_ons), vjust=1.6, color="white",
            position = position_stack(vjust = 0.5), size=3.5) +
    geom_line(data=bdat, aes(x=Period, y=max_load), group=1, size=2, color='grey') +
    geom_point(data=bdat, aes(x=Period, y=max_load, size=4, fill='black')) +
    theme_minimal() + 
    theme(legend.position = "none") +
    ggtitle(titleb) +
    ylab("Passengers")
  print(b)
})
```

Row
----------------------------
### Passengers per Vehicle Hour

```{r message=FALSE, warning=FALSE}
renderPlot({
  cdat <- subset(data, Route==input$filterA)
  titlec <- sprintf("Mean Passengers per Vehicle Hour - Route %s", input$filterA)
  c <- ggplot(data=cdat, aes(x=Period, y=mean_ppvh, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_ppvh), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
    theme_minimal() + 
    theme(legend.position = "none") +
    ggtitle(titlec) +
    ylab("Passengers")
  print(c)
})
```

### Passenger Miles

```{r message=FALSE, warning=FALSE}
renderPlot({
  ddat <- subset(data, Route==input$filterA)
  titled <- sprintf("Passenger Miles per Trip - Route %s", input$filterA)
  d <- ggplot(data=ddat, aes(x=Period, y=mean_pm, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_pm), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
    theme_minimal() + 
    theme(legend.position = "none") +
    ggtitle(titled) +
    ylab("Passenger Miles per Trip")
  print(d)
})
```

Performance {data-icon="fa-clipboard-check"}
==================

Row
-----------------------------

### OTP: November 2019

```{r message=FALSE, warning=FALSE}
renderGauge({
  datag1 <- subset(OTP, Route==input$filterA)
  titleg1 <- sprintf("Route %s", input$filterA)
  rate1 <- round(((datag1$NOV19)*100),0)
  gauge(rate1, min = 0, max = 100, symbol = '%', label=titleg1, gaugeSectors(
    success = c(79, 100), warning = c(70, 78), danger = c(0, 69)))
})
```

### OTP: December 2019

```{r message=FALSE, warning=FALSE}
renderGauge({
  datag2 <- subset(OTP, Route==input$filterA)
  titleg2 <- sprintf("Route %s", input$filterA)
  rate2 <- round(((datag2$DEC19)*100),0)
  gauge(rate2, min = 0, max = 100, symbol = '%', label=titleg2, gaugeSectors(
    success = c(79, 100), warning = c(70, 78), danger = c(0, 69)))
})
```

### OTP: January 2020
```{r message=FALSE, warning=FALSE}
renderGauge({
  datag3 <- subset(OTP, Route==input$filterA)
  titleg3 <- sprintf("Route %s", input$filterA)
  rate3 <- round(((datag3$JAN20)*100),0)
  gauge(rate3, min = 0, max = 100, symbol = '%', label=titleg3, gaugeSectors(
    success = c(79, 100), warning = c(70, 78), danger = c(0, 69)))
})
```

### OTP: February 2020
```{r message=FALSE, warning=FALSE}
renderGauge({
  datag4 <- subset(OTP, Route==input$filterA)
  titleg4 <- sprintf("Route %s", input$filterA)
  rate4 <- round(((datag4$FEB20)*100),0)
  gauge(rate4, min = 0, max = 100, symbol = '%', label=titleg4, gaugeSectors(
    success = c(79, 100), warning = c(70, 78), danger = c(0, 69)))
})
```

### OTP: March 2020
```{r message=FALSE, warning=FALSE}
renderGauge({
  datag5 <- subset(OTP, Route==input$filterA)
  titleg5 <- sprintf("Route %s", input$filterA)
  rate5 <- round(((datag5$MAR20)*100),0)
  gauge(rate5, min = 0, max = 100, symbol = '%', label=titleg5, gaugeSectors(
    success = c(79, 100), warning = c(70, 78), danger = c(0, 69)))
})
```

Row
-----------------------------

### Miles per Hour

```{r message=FALSE, warning=FALSE}
renderPlot({
  jdat <- subset(data, Route==input$filterA)
  j <- ggplot(data=jdat, aes(x=Period, y=mean_mph, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_mph), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
    ggtitle("Mean Miles per Hour") +
    theme_minimal() + 
    theme(legend.position = "none") +
    ylab("Miles per Hour")
  print(j)
})
```

----------------------------
### Trip Duration

```{r message=FALSE, warning=FALSE}
renderPlot({
  kdat <- subset(data, Route==input$filterA)
  k <- ggplot(data=kdat, aes(x=Period, y=mean_duration, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_duration), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
    ggtitle("Mean Trip Duration") +
    theme_minimal() + 
    theme(legend.position = "none") +
    ylab("Minutes")
  print(k)
})
```

----------------------------
### Average Passengers per Mile

```{r message=FALSE, warning=FALSE}
renderPlot({
  ldat <- subset(data, Route==input$filterA)
  l <- ggplot(data=ldat, aes(x=Period, y=mean_ppm, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_ppm), vjust=1.6, color="white", position = position_dodge(0.9), size=3.5) +
    ggtitle("Mean Passengers per Mile") +
    theme_minimal() + 
    theme(legend.position = "none") +
    ylab("Passengers per Mile")
  print(l)
})
```

Operations {data-icon="fa-bus"}
==================

Row
-----------------------------

### OTP

```{r message=FALSE, warning=FALSE}
renderPlot({
  mdat <- subset(OTP_hist, Route==input$filterA)
  mdat$value <- round(((mdat$value)*100),0)
  m <- ggplot(mdat, aes(fill=category, y=value, x=Date)) + 
    geom_bar(position="stack", stat="identity") +
    geom_text(aes(label=value), color="white", size = 3, position = position_stack(vjust = 0.5)) +
    theme_minimal() + 
    ggtitle("12-month On-Time Performance") +
    ylab("Percentage") +
    scale_fill_manual(name= " ", values = c("#00BA38", "#F8766D", "#619CFF"), labels = c("% Early", "% Late", "% On-Time")) +
    scale_x_discrete(breaks=c("2019-04", "2019-06", "2019-08", "2019-10", "2019-12", "2020-02"))
  print(m)
})
```

### Passengers per Vehicle Hour

```{r message=FALSE, warning=FALSE}
renderPlot({
  n1dat <- subset(data, Route==input$filterA)
  n1 <- ggplot(data=n1dat, aes(x=Period, y=mean_ppvh, fill=Period)) +
    geom_bar(stat='identity') +  
    geom_text(aes(label=mean_ppvh), vjust=1.6, color="white", position = position_dodge(0.9), size=3.5) +
    theme_minimal() + 
    theme(legend.position = "none") +
    ggtitle("Average Passengers per Vehicle Hour") +
    ylab("Passengers")
  print(n1)
})
```

Comparison {data-icon="fa-chart-line"}
==================
Hover over a data point to learn more.

Row
-----------------------------

### Cost per Passenger
```{r message=FALSE, warning=FALSE}
output$oout <- renderPlotly({
  highlighto <- subset(opstats, Route==input$filterA)
  titleo <- sprintf("Cost per Passenger - Route %s", input$filterA)
  o <- ggplot(data=opstats, aes(x=avg_wkdy_pax, y=cost_pp, name=Route, color=Classification)) +
    geom_point() +
    geom_point(data=highlighto, aes(x=avg_wkdy_pax, y=cost_pp), color='black', size=3) +
    theme_minimal() + 
    ggtitle(titleo) +
    ylab("Vehicles") +
    xlab("Average Weekday Passengers")
  buildo <- plotly_build(o)
  buildo
})
plotlyOutput('oout') 
```

### Peak Vehicles
```{r message=FALSE, warning=FALSE}
output$pout <- renderPlotly({
  highlightp <- subset(opstats, Route==input$filterA)
  titlep <- sprintf("Peak Vehicles - Route %s", input$filterA)
  p <- ggplot(data=opstats, aes(x=avg_wkdy_pax, y=peakv, name=Route, color=Classification)) +
    geom_point() +
    geom_point(data=highlightp, aes(x=avg_wkdy_pax, y=peakv), color='black', size=3) +
    theme_minimal() + 
    ggtitle(titlep) +
    ylab("Vehicles") +
    xlab("Average Weekday Passengers")
  buildp <- plotly_build(p)
  buildp
})
plotlyOutput('pout')
```

Row
-----------------------------

### Passengers per Vehicle Hour
```{r message=FALSE, warning=FALSE}
output$qout <- renderPlotly({
  highlightq <- subset(opstats, Route==input$filterA)
  titleq <- sprintf("Passengers per Revenue Hour - Route %s", input$filterA)
  q <- ggplot(data=opstats, aes(x=avg_wkdy_pax, y=pprh, name=Route, color=Classification)) +
    geom_point() +
    geom_point(data=highlightq, aes(x=avg_wkdy_pax, y=pprh), color='black', size=3) +
    theme_minimal() + 
    ggtitle(titleq) +
    ylab("Passengers") +
    xlab("Average Weekday Passengers")
  buildq <- plotly_build(q)
  buildq
})
plotlyOutput('qout')
```

### Vehicle Miles
```{r message=FALSE, warning=FALSE, cache=TRUE}
output$rout <- renderPlotly({
  highlightr <- subset(opstats, Route==input$filterA)
  titler <- sprintf("Vehicle Miles - Route %s", input$filterA)
  r <- ggplot(data=opstats, aes(x=avg_wkdy_pax, y=vm, name=Route, color=Classification)) +
    geom_point() +
    geom_point(data=highlightr, aes(x=avg_wkdy_pax, y=vm), color='black', size=3) +
    theme_minimal() + 
    ggtitle(titler) +
    ylab("Vehicle Miles") +
    xlab("Average Weekday Passengers")
  buildr <- plotly_build(r)
  buildr
})
plotlyOutput('rout')
```

Map {data-icon="fa-map"}
==================
```{r message=FALSE, warning=FALSE}
renderLeaflet({
  mapselection <- input$filterA
  mapselection[mapselection=='LUCY GO'] <- 'LUCYGO'
  maproute <- subset(routelines, Route==mapselection)
  ridership <- subset(ridership, Route==mapselection)
  ridershipa <- subset(ridership, Direction == "Eastbound" | Direction=="Northbound" | Direction=="Inbound" | Direction=="Loop")
  ridershipb <- subset(ridership, Direction == "Westbound" | Direction=="Southbound" | Direction=="Outbound")
  routeno <- sprintf("Route %s", input$filterA)
  routeA <- sprintf("Ridership %s", ridershipa$Direction)
  routeB <- sprintf("Ridership %s", ridershipb$Direction)
  
  s <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron, group =  "Basemap") %>%
  addPolylines(data = maproute,
               weight = 3, 
               group = routeno) %>%
  addCircles(data = ridershipa,
             weight = 1, 
             color = 'royalblue',
             radius = ridershipa$Weekday_Boards+ridershipa$Weekday_Leaves, 
             popup = paste("Stop:", ridershipa$Stop_Name, "<br>",
                           "Direction:", ridershipa$Direction, "<br>",
                           "Weekday Boards:", ridershipa$Weekday_Boards,"<br>",
                           "Weekday Leaves:", ridershipa$Weekday_Leaves, "<br>"),
             group = routeA
  )%>%
  addCircles(data = ridershipb,
             weight = 1, 
             color = '#ff9900',
             radius = ridershipb$Weekday_Boards+ridershipb$Weekday_Leaves,
             popup = paste("Stop:", ridershipb$Stop_Name, "<br>",
                           "Direction:", ridershipb$Direction, "<br>",
                           "Weekday Boards:", ridershipb$Weekday_Boards, "<br>",
                           "Weekday Leaves:", ridershipb$Weekday_Leaves, "<br>"), 
             group = routeB)%>%
  addLayersControl(
    overlayGroups = c("Basemap", routeno, routeA, routeB),
    options = layersControlOptions(collapsed = FALSE)
  )%>%
  addLegend(
    position = "bottomright",
    colors = c("royalblue", "#ff9900"), 
    labels = c("RouteA", "RouteB"))
  s
})

```
